---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
library(badger)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# bpcs - A package for Bayesian Paired Comparison analysis with Stan

<!-- badges: start -->
`r badge_license('MIT', color='yellow')`
`r badge_last_commit("davidissamattos/bpcs")`
`r badge_lifecycle("experimental", "blue")`
`r badge_code_size("davidissamattos/bpcs")`
[![codecov](https://codecov.io/gh/davidissamattos/bpc/branch/master/graph/badge.svg?token=6RTC7768CL)](undefined)

<!-- badges: end -->

The `bpcs` package performs Bayesian estimation of Paired Comparison models utilizing Stan. 


## Installation

For the `bpcs` package to work, we rely upon the Stan software and the `rstan` package. 

* For general installation of Stan and RStan see: https://github.com/stan-dev/rstan/wiki/RStan-Getting-Started
* For macOS: https://github.com/stan-dev/rstan/wiki/Installing-RStan-from-source-on-a-Mac
* For Linux: https://github.com/stan-dev/rstan/wiki/Installing-RStan-on-Linux
* For Windows: https://github.com/stan-dev/rstan/wiki/Installing-RStan-from-source-on-Windows 

At the moment we are not available in CRAN. To install the bpc package install directly from our Github repository

```{r eval=FALSE, echo=T}
remotes::install_github('davidissamattos/bpcs')
```

After install we load the package with:
```{r}
library(bpcs)
```


## Example

The main function of the package is the `bpc` function. This function requires a specific type of data frame that contains:

* Two columns containing the name of the contestants in the paired comparison
* Two columns containing the score of each player OR 
* one column containing the result of the match (0 if player0 won, 1 if player1 won, -1 if it was a tie)

We will utilize a prepared version of the known dataset available in Agresti 2002. The same dataset can also be found in other packages such as the `BradleyTerry2`. The dataset can be seen below and is available as `data(citations_agresti)`:

```{r}
knitr::kable(citations_agresti)
```

Based on the scores of each contestant, the `bpc` function computes automatically who won the contest. Alternatively, you can provide a vector of who won if that is already available (for more information see `?bpc`.

For the simple Bradley Terry Model we specify the model type as `'bradleyterry'` . If there are ties in the data we can use a different model (see vignette: TODO), or we can solve it randomly with the `'random'` option. Since we provide the scores we need to say who won `'higher'` if higher score wins or `'lower'` if lower score wins. 

```{r}
m1<-bpc(data=citations_agresti,
                  player0 = 'journal1',
                  player1 = 'journal2',
                  player0_score = 'score1',
                  player1_score = 'score2',
                  model_type='bradleyterry',
                  solve_ties='random',
                  win_score = 'higher')
```

If `rstan` is available and correctly working this function should sample the posterior distribution and create a `bpc` object.

To see a summary of the results we can run:

```{r}
summary(m1)
```


## Vignettes

This package provides a series of small and self contained vignettes that exemplify the use of each model. In the vignettes, we also provide examples of code for data transformation, tables and plots.

Below we list all our vignettes:

* Getting Started: TODO
* Handling ties: TODO
* Bradley-Terry with random effects: TODO
