---
output: github_document
bibliography: bibliography.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
library(badger)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# bpcs - A package for Bayesian Paired Comparison analysis with Stan

<!-- badges: start -->
`r badge_license('MIT', color='yellow')`
`r badge_last_commit("davidissamattos/bpcs")`
`r badge_lifecycle("experimental", "blue")`
`r badge_code_size("davidissamattos/bpcs")`

[![Codecov test coverage](https://codecov.io/gh/davidissamattos/bpcs/branch/master/graph/badge.svg)](https://codecov.io/gh/davidissamattos/bpcs?branch=master)
<!-- badges: end -->

The `bpcs` package performs Bayesian estimation of Paired Comparison models utilizing Stan. 

Package documentation and vignette articles can be found at: https://davidissamattos.github.io/bpcs/


## Installation

For the `bpcs` package to work, we rely upon the Stan software and the `rstan` package [@rstan]. 

* For general installation of Stan and RStan see: https://github.com/stan-dev/rstan/wiki/RStan-Getting-Started
* For macOS: https://github.com/stan-dev/rstan/wiki/Installing-RStan-from-source-on-a-Mac
* For Linux: https://github.com/stan-dev/rstan/wiki/Installing-RStan-on-Linux
* For Windows: https://github.com/stan-dev/rstan/wiki/Installing-RStan-from-source-on-Windows 

At the moment we are not available in CRAN. To install the bpc package install directly from our Github repository

```{r eval=FALSE, echo=T}
remotes::install_github('davidissamattos/bpcs')
```

After install we load the package with:
```{r}
library(bpcs)
```


## Minimal example

The main function of the package is the `bpc` function. This function requires a specific type of data frame that contains:

* Two columns containing the name of the contestants in the paired comparison
* Two columns containing the score of each player OR 
* one column containing the result of the match (0 if player0 won, 1 if player1 won, -1 if it was a tie)

We will utilize a prepared version of the citations dataset available [@agresti2003categorical]. The same dataset can also be found in other packages that implement the Bradley-Terry model such as the `BradleyTerry2` [@turner2012bradley]. The dataset can be seen below and is available as `data(citations_agresti)`:

```{r}
knitr::kable(citations_agresti)
```

Based on the scores of each contestant, the `bpc` function computes automatically who won the contest. Alternatively, you can provide a vector of who won if that is already available (for more information see `?bpc`.


For the simple Bradley Terry Model we specify the model type as `'bradleyterry'`. Here we hide the MCMC sampler chain messages for simplicity. 

```{r}
m1<-bpc(data=citations_agresti,
                  player0 = 'journal1',
                  player1 = 'journal2',
                  player0_score = 'score1',
                  player1_score = 'score2',
                  model_type='bradleyterry',
                  solve_ties='random',
                  win_score = 'higher',
        show_chain_messages = F)
```

If `rstan` is available and correctly working this function should sample the posterior distribution and create a `bpc` object.

To see a summary of the results we can run:

```{r}
summary(m1)
```


## Vignettes

This package provides a series of small and self contained vignettes that exemplify the use of each model. In the vignettes, we also provide examples of code for data transformation, tables and plots.

Below we list all our vignettes:

* Getting Started: TODO
* Handling ties: TODO
* Bradley-Terry with random effects: TODO
